\documentclass[12pt]{article}
\usepackage{exercises/handwritten/handout}

\usepackage{amssymb,mathrsfs, amsmath,amsfonts}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{braket}
\usepackage{hyperref}
%%
%%
\newcommand{\Blank}{\mbox{\hskip 4pt\vrule width 1in depth 2pt}\vrule width 0pt height 2.0em}
\newcommand{\NameBlank}{\mbox{\hskip 4pt\vrule width 2.5in depth 2pt}\vrule width 0pt height 2.0em}
%%
%% Leave at least #1 space, default to what is below
%%
\def\DefaultSpace{1in}
\newcommand{\LeaveSpace}[1][\DefaultSpace]{%
\vskip #1 plus 1fil\relax\hbox to 0pt{\hss} %
}

\begin{document}

\assignment{Problem Set 0}

\noindent Name:\NameBlank{}   Student ID:\NameBlank{} \newline

\begin{quote}
    As for all written assignments, you must complete the work on these pages, and upload your completed pages to GradeScope.  Do not simply submit a page with the answers, as we cannot grade your work that way.  GradeScope will allow you to submit separate images for reach response, but it's on you to submit those properly.

    Work submitted improperly will receive no credit.
\end{quote}

\begin{enumerate}[font=\bfseries]
    \item (4 points) Recall we say a computation (or function) is \emph{reversible} if it is always possible to uniquely recover the input, given the output. Assume $x \in \mathbb{R}$.
    \begin{itemize}
        \item If $f : x \mapsto x^2$, is $f$ a reversible function? \Blank{}
        \item If $f : x \mapsto x^3$, is $f$ a reversible function? \Blank{}
        \item If $f : x \mapsto (x^2,\textrm{sign}(x))$, is $f$ a reversible function? \Blank{}
        \item If $f : x \mapsto 0$, is $f$ a reversible function? \Blank{}
    \end{itemize}
    \item (2 points) What quadrant of the complex unit circle is $e^{i\frac{3\pi}{4}}$ in?\Blank{}
    \item (4 points) Suppose $\ket{a} = \begin{pmatrix}5 \\ i \end{pmatrix}$ and $\ket{b} = \begin{pmatrix}2 \\ 3 \end{pmatrix}$.\begin{itemize}
        \item Solve $\braket{a|b} = $ \Blank{} Solve $\ket{b}\bra{a} = $ \Blank{}
        \item Solve $\braket{b|b} = $ \Blank{} Solve $|\ket{b}|^2 = $ \Blank{}
    \end{itemize}
    \item (2 points) Why did you decide to take this course? \newline
    
\end{enumerate}



\end{document}
