\documentclass[12pt]{article}
\usepackage{amssymb,mathrsfs, amsmath,amsfonts}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{braket}

\title{Problem Set 0 Solutions}
\author{CSE 468}
\date{\today}

\begin{document}

\maketitle

\begin{enumerate}[font=\bfseries]
    \item Yes. $f(x)$ has the same number of inputs and outputs (namely 1). Furthermore, we can use the inverse $f^{-1}(x) = x^{\frac{1}{3}}$, to obtain the input.
    \item No. A set may contain many elements while the max function outputs a single number. Thus, it is impossible to determine the whole set given just the maximum value of the set, meaning max is not a reversible function.
    \item No. SHA-256 uses a ``one-way compression function" which takes two inputs and returns a single output.
    \item Recall that we can think of theta as the distance we traverse around the complex unit circle starting from $(1,0)$. By tracing this out on a unit circle, it can be seen that $e^{i\frac{3\pi}{4}}$ is in quadrant 2.
    Alternatively, we can use $e^{i\frac{3\pi}{4}} = \cos{\frac{3\pi}{4}}+i\sin{\frac{3\pi}{4}} = -\frac{1}{\sqrt{2}}+i\frac{1}{\sqrt{2}}$.
    The real component, $-\frac{1}{\sqrt{2}}$, tells us we're on the left side of the unit circle. The imaginary coefficient, $\frac{1}{\sqrt{2}}$, puts us in the top half. The intersection of these pieces of information yields quadrant 2.
    \item 10 - 3$i$
    \item Acceptable answers:
        \begin{itemize}
            \item Because quantum computing is cool
            \item Because Ron, David, and Will are cool
            \item Etc.
        \end{itemize}
\end{enumerate}



\end{document}
