\lecture[0]{Introduction}{lec:intro}

\title{Introduction}
\subtitle{Getting started}

\date{\today}

\begin{frame}
\maketitle
\end{frame}
\section{Instructor}
\begin{frame}{Your professor}{Brief bio}
\begin{itemize}
    \item Ron Cytron
    \begin{itemize}
        \item You can call me \emph{Ron}
        \item Last name pronounced sit'-run
    \end{itemize}
    \item Undergrad at Rice University
    \item Graduate MS and PhD at University of Illinois
    \item Primary research interests
    \begin{itemize}
        \item Programming languages and compilers
        \item Runtime systems
        \item Computer architecture
    \end{itemize}
    \item So how did I get to Quantum Computing?
\end{itemize}
    
\end{frame}

\section{Background}
\begin{frame}{Curiosity}{And I had a lot of help $\ldots$}
\begin{itemize}
    \item Walter Buhro -- student in CSE 131 but also physics major
    \item Nathan Mester -- we studied this material over a summer
    \item Then I got serious
    \begin{itemize}
        \item Arthur Rattew
        \item Collin Szczepanski
        \item Finn Voichick
    \end{itemize}
\end{itemize}
    
\end{frame}

\begin{frame}{Quantum Computing}{Inherently multidisciplinary}
\begin{tabular}{cc}
\begin{tikzpicture}[scale=0.6]
%  \begin{scope}
    \fill<1>[color=BrickRed]   ( 90:1.2) circle (2);
    \draw[color=BrickRed]   ( 90:1.2) circle (2);
    \fill<2>[color=OliveGreen] (210:1.2) circle (2);
    \draw[color=OliveGreen] (210:1.2) circle (2);
    \fill<3>[color=NavyBlue]  (330:1.2) circle (2);
    \draw[color=NavyBlue]  (330:1.2) circle (2);
%  \end{scope}
  \node<1-> at ( 90:2)    {Physics};
  \node<2-> at ( 210:2)   {Math};
  \node<3-> at ( 330:2)  {CS};
  \node<4-> [font=\Large] {QC};
\end{tikzpicture}
&
\begin{minipage}[b]{0.5\textwidth}
\begin{itemize}
    \item<1->{\textcolor<1>{BrickRed}{Physics is the study of the observable universe}}
    \item<2->{\textcolor<2>{OliveGreen}{Math is the study of logic and reasoning}}
    \item<3->{\textcolor<3>{NavyBlue}{Computer science solves problems using computation}}
\end{itemize}

\end{minipage}
\end{tabular}

\bigskip

\only<4>{

\bigskip

Quantum Computing combines all of these disciplines}
\only<1>{\ColorR{We study physics so as to understand and believe in quantum behavior}}
\only<2>{\ColorG{Quantum values and gates can be represented using linear algebra}}
\only<3>{\ColorB{Computer science lets us reason about the problems we can solve}}
\end{frame}

\begin{frame}{Prerequisites}
\begin{description}
    \item[Linear and Matrix Algebra] We use this heavily.  You must be familiar with row and column vectors, matrix multiplication, linear systems of equations.
    
    Example:  Math 309
    \item[Probability] Quantum computations are inherently probabilistic.  You need to have some background in probability, but it doesn't have to be extensive.
    
    Example:  ESE 326 or Math 3200
    \item[Circuits]  We will specify quantum computations using circuit diagrams.  Some familiarity with Boolean algebra and circuit elements depicting Boolean operations is needed.
    
    Example:  CSE 132 or CSE 260M
\end{description}
\end{frame}
\begin{frame}{Goals}
\begin{itemize}
    \item Understand and be able to construct quantum circuits to solve a problem.
    \item Sufficient background to take quantum information courses in ESE.
    \item Sufficient background to take foundational courses in Math.
\end{itemize}
\end{frame}

\section{Computation}

\begin{frame}{Computer scientists are cuckoos}{We steal ideas from other disciplines to compute things}

We can compute things using
\begin{itemize}
    \item Steam
    \item Semiconductor physics
    \item Water
    \item DNA
    \item \alert{Gravity}
    \item Quantum effects
\end{itemize}
    
\end{frame}

\begin{frame}{Using gravity to compute square roots}{We steal ideas from other disciplines to compute things}
\only<1->{How can we use gravity to compute square roots?}

\only<2-3>{We know from physics}
\only<3->{
\begin{eqnarray*}
\only<3>{\ColorR{d} & = & \frac{1}{2} a \ColorG{t}^{2} \\}
\only<4->{\ColorG{t} & = & \sqrt{\frac{2\ColorR{d}}{a}}}
\end{eqnarray*}}
\only<5->{To compute $\sqrt{n}$, we just need to drop a ball from height $\ColorR{d}=\frac{an}{2}$ and measure the time for the ball to hit the ground.}
\only<6->{

\medskip

Example:
\begin{eqnarray*}
a & = & 9.8 \mbox{ m/$s^2$} \\
n & = & 81 \\
\ColorR{d} & = & an / 2 = 369.9 \mbox{ meters}
\end{eqnarray*}
Takes 9 seconds to drop}
\end{frame}