\begin{frame}{Some notes about online learning for this course}{There are challenges but also some opportunities}

\Vskip{-3.5em}
\only<1-6>{\begin{itemize}[<+->]
    \item We all wish we could be in person.  I was particularly looking forward to studying this material with you in person.  It's difficult to impart or sense amazement on \href{https://en.wikipedia.org/wiki/Zoom_Video_Communications}{zoom}.
    \item While we are forced to study together on zoom, I ask the following of you:
    \begin{itemize}
        \item Your participation grade is partly based on attendance. I will have a form you use to sign in each class.  Beyond that part of your grade, your success in this class is likely improved by just \alert{showing up}.
        \item Unless it would be improprietous, \alert{please leave your video on} so I can see your face.  Teaching on zoom otherwise is like teaching in a classroom where students have bags over their heads. 
        \item I strongly encourage you to exercise curiosity and ask questions. If questions show up in chat, I may not see them while screen sharing.  \alert{Please interrupt me if you see a question that I haven't seen.}  Also feel free to \alert{ask questions out loud}.
    \end{itemize}
    \item It is challenging to give exams in an online course.  Please take note of my \href{https://wustl.instructure.com/courses/80191}{course policies}, especially regarding exams.  I will go over those today. 
\end{itemize}}%
\only<7>{%
\begin{itemize}
    \item However, there are some advantages.  One is that you can annotate the screen while I am sharing, to ask a question or to participate in a discussion.  Look for \texttt{View Options}$\ldots$\texttt{Annotate}$\ldots$\texttt{Draw}
    \item Let's practice that now.  In the spirit of \href{https://admission.rice.edu/blog/in-the-know/your-application-checklist}{Rice's application Box}, please draw or write something \ColorOne{in the square below} now:
\end{itemize}
\ColorOne{\fbox{\hbox to 0.98\textwidth{\hss}\vrule width 0pt depth 1.5in}}}%
\only<8->{%
\begin{itemize}
    \item When we return, hopefully, to in-person instruction, we have been asked to continue to post online recordings of lectures.
    \item I have agreed to do that, based on a quick review of literature that shows outcomes improve when such material is available.
    \item However, captures of live lectures are not the best way to present material online.
    \item You are both encouraged and incentivized (class participation credit) to attend in-person whenever possible.
    \item However, if you feel ill or are concerned about communicating COVID or other illness to yourself or others, please do not attend.
    \item A form will be provided so you can receive credit for participation whether you are present or (with good reason) absent.
\end{itemize}
}
    
\end{frame}
