\documentclass[12pt]{article}
\usepackage{amssymb,mathrsfs, amsmath,amsfonts}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{braket}

\title{PS5 Grading Guide / Rubric}
\author{CSE 468}
\date{\today}

\begin{document}

\maketitle

Assignment: 16 points total
\begin{enumerate}[font=\bfseries]
    \item 2 points - What is the state of Bob's qubits?
        \begin{itemize}
            \item 2 points - Correct answer
            \item -1 point for small error
        \end{itemize}
    \item 2 points - What two outcomes might he see for his $\mathbf{Z}$ measurement?
        \begin{itemize}
            \item 1 point for each correct answer
        \end{itemize}
    \item 2 points - What two outcomes might he see for his $\mathbf{X}$ measurement?
        \begin{itemize}
            \item 1 point for each correct answer
        \end{itemize}
    \item 2 points - Are all combinations of outcomes possible?
        \begin{itemize}
            \item 1 point for recognizing not all outcomes are possible
            \item 1 point for recognizing the correct possible combinations
        \end{itemize}
    \item 2 points - What two possible states could Bob see when he measures after Alice?
        \begin{itemize}
            \item 1 point for each correct answer
        \end{itemize}
    \item 3 points - What result does Bob report for each of his squares?
        \begin{itemize}
            \item 1 point for each correct answer
        \end{itemize}
    \item 1 point - What value must Alice have reported for her first square?
        \begin{itemize}
            \item 1 point for correct answer
        \end{itemize}
    \item 2 points - What information on Slide 14 assures Bob and Alice see the same result for common square?
        \begin{itemize}
            \item 2 points for correct response
        \end{itemize}
\end{enumerate}


\end{document}