\documentclass[12pt]{article}
\usepackage{amssymb,mathrsfs, amsmath,amsfonts}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{braket}

\title{Problem Set 2}
\author{CSE 468}
\date{May 2021}

\begin{document}

\maketitle

\noindent \textbf{Note:} Pauli matrix type questions

\begin{enumerate}[font=\bfseries]
    \item Given $\ket{\psi} = \alpha\ket{0} + \beta\ket{1}$,
    \begin{enumerate}
        \item What is the state of $\ket{\psi_0} = \mathbf{X}\mathbf{Y}\mathbf{Z}\ket{\psi}$?
        \item Describe the above operation as rotations on the Bloch sphere.
        \item What is the state of $\ket{\psi_1} = \mathbf{Z}\mathbf{Y}\mathbf{X}\ket{\psi}$?
        \item Does $\ket{\psi_0} = \ket{\psi_1}$?
        \item Is the probability of measuring $\ket{0}$ when we measure $\ket{\psi_0}$ equal to the probability of measuring $\ket{0}$ when we measure $\ket{\psi_1}$?
    \end{enumerate}
    \item Give a unitary matrix that when applied to $\ket{0}$ (in the standard basis) results in a $\frac{3}{4}$ chance of measuring $\ket{1}$ in the $\mathbf{Y}$ basis. 
    \item Consider $\ket{\psi} = \mathbf{X}^a\mathbf{Y}^b\mathbf{Z}^c\ket{0}$.
    \begin{enumerate}
        \item If $a,c$ are even and $b$ is odd, what is the probability of observing $\ket{0}$ when we measure $\ket{\psi}$?
        \item If $a,b,c$ are all even what is the chance of measuring $\ket{1}$ in the $\mathbf{X}$ basis?
    \end{enumerate}
    \item If we begin at $\ket{0}$ on the Bloch sphere and can use any of the Pauli gates, how many unique locations can we travel to on the Bloch sphere?
    \item If we now also allow the use of the Hadamard gate, how many unique locations can we travel to on the Bloch sphere?
    \item Recall two points on the Bloch sphere are antipodal if they determine a line that passes through the center of the sphere. Recall that in polar coordinates, a quantum state $\ket{\psi}$ can be specified on the Bloch sphere using two coordinates, $\theta$ and $\phi$:
    \[\ket{\psi} = \cos(\frac{\theta}{2})\ket{0}
    + e^{i\phi}\sin(\frac{\theta}{2})\ket{1}\]
    \begin{enumerate}
        \item All points on the Bloch sphere denote unique quantum states except which two points?
        \item The number of quantum states on the Bloch sphere is \emph{countably} infinite or \emph{uncountably} infinite?
        \item Given a state $\ket{\psi}$ defined as above, define its antipodal state $\ket{\psi'}$
        \item Prove that $\ket{\psi}$ and $\ket{\psi'}$ are orthogonal by showing their inner product is 0.
    \end{enumerate}
    \item Your friend, Albert Einstein, has challenged you to a quantum game. The game is played as follows: First, Einstein gives you some state $\ket{\psi}$ and tells you where it is located on the Bloch sphere. You then can apply any quantum gates you wish to $\ket{\psi}$. Before you make a measurement you tell Einstein what state you expect to observe. If you are right Einstein gives you \$100. If you are wrong Einstein zaps you with his many-worlds invention. Do you agree to play Einstein's game? Why or why not?
    \item Consider the quantum states
        \[\ket{+y} = \frac{1}{\sqrt{2}}\ket{0} + \frac{i}{\sqrt{2}}\ket{1}\]
        and
        \[\ket{-y} = \frac{-i}{\sqrt{2}}\ket{0} - \frac{1}{\sqrt{2}}\ket{1}\]
        What quantum gate (unitary matrix) $U$ produces those states given the input qubits $\ket{0}$ and $\ket{1}$, respectively?
        \begin{enumerate}
            \item Define $U$.
            \item Prove $U$ is unitary.
            \item Write $U$ as the outer product of $\ket{0}$, $\ket{1}$, $\ket{+y}$, and $\ket{-y}$.
            \item More generally, what do the columns of such as 2 × 2 matrix represent with respect to their inputs in the standard basis ($\ket{0}$ and $\ket{1}$)?
        \end{enumerate}
\end{enumerate}



\end{document}