\documentclass[12pt]{article}
\usepackage{amssymb,mathrsfs, amsmath,amsfonts}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{braket}
\graphicspath{ {./ps4-assets/}{./exercises/handwritten/ps4-assets/} }


\title{Problem Set 5}
\author{CSE 468}
\date{May 2021}

\begin{document}

\maketitle



\begin{enumerate}[font=\bfseries]
    \item Note: Key distribution, CHSH, Mermin peres
    \item Could copy and change a few values in table on slide 8 in lecture 11. Have students calculate the shared key. Or have students find the line in the table that indicates Eve is present. How many key bits should they publish to be x confident?
    \item Suppose you and your friend Bob are trying to decide between using the E91 and BB84 protocols? What factors should you take into account? When would you prefer one protocol over the other?
    \item What is the primary property of quantum mechanics that enables the BB84 protocol?
    \item What is the primary property of quantum mechanics that enables the E91 protocol?
    \item B92 protocol question (basically BB84 with only two orthogonal states). Give brief description of protocol. Ask about how to detect Eve and probability bound.
    \item I will provide an (insecure) protocol. Students must identify the problem in the protocol (prblm will likely be being unable to detect Eve)
    \item Consider the BB84 protocol. Why must Alice wait until after Bob has measured all photons before publishing the bases she measured in?
    \item 3 person CHSH? Not exactly sure how this would work
    \item More CHSH and MP questions
\end{enumerate}



\end{document}