\documentclass[12pt]{article}
\usepackage{amssymb,mathrsfs, amsmath,amsfonts}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{braket}

\title{Creating Programming Assignments}
\author{CSE 468}
\date{May 2021}

\begin{document}

\maketitle

\noindent \textbf{Note:} You must have otter-grader, pandoc, and some form of TeX (will add more clarification on these requirements soon). Also thanks to Will Sepesi for some of these notes.

\begin{enumerate}[font=\bfseries]
    \item Create a programming assignment on Gradescope. Configure settings as desired.
    \item Create a homework assignment on Gradescope. Use a placeholder PDF as the template. Set the Submissions option to Instructor. (Should I set Submission Type to Variable or Fixed length? Need to investigate further). Ensure template visibility is set to false. Configure other settings as desired.
    \item Clone sample.ipynb. Fill in the token, course\_id, and assignment\_id fields. The course\_id and assignment\_id can be found on Gradescope. (ex. https://www.gradescope.com/courses/280928/assignments/1489113/) 280928 is the course\_id and 1489113 is the assignment\_id. The assignment\_id should be for the written homework assignment. To get your token:
    \begin{enumerate}
        \item Open a python shell
        \item from otter.generate.token import APIClient
        \item print(APIClient.get\_token())
    \end{enumerate}
    \item Run all the cells in your notebook and save
    \item Run otter assign yourAssignment.ipynb destName --v1
    \item Add the autograder.zip to the programming assignment on Gradescope
    \item In the written homework assignment on Gradescope replace your placeholder PDF with the generated template in the autograder.zip. Create the problems and problem regions as desired. (Add clarification here)
    \item Push to students (best way to do this?)
\end{enumerate}

Will add student instructions TBD

\end{document}