\documentclass[12pt]{article}
\usepackage{exercises/handwritten/handout}

\usepackage{amssymb,mathrsfs, amsmath,amsfonts}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{braket}
\input{quantummacros.tex}
%%
%%

%%
%% Leave at least #1 space, default to what is below
%%


\begin{document}

\assignment{Problem Set 330}

\begin{itemize}
    \item Organize yourselves in groups of up to 4 people.
    \item Access the \href{https://docs.google.com/document/d/1gwp3NOt7oRPQ2ymmrp5rFvVzPTFJ6iV4FJLHPvpXlO8/edit?usp=sharing}{Research Questions} document and modify the document under the appropriate research question to show what you will study.
    \item The starting notebook is on Canvas, called \texttt{BVSolved.ipynb}, in Module 33 for Errors and Error Detection.
    \item This assignment is worth 100 points but you get those points just by completing the work.
\end{itemize}

\begin{enumerate}[font=\bfseries]

\item Organize yourselves in groups up to four people.  Your people are (names and student IDs please):
\begin{itemize}
    \item \Blank[4in]{}
    \item \Blank[4in]{}
    \item \Blank[4in]{}
    \item \Blank[4in]{}
\end{itemize}
\item What questions are you working on from the research document?
\textbf{Be sure you modified the document so others can see what you are doing.}
\LeaveSpace[3in]{}
\item Describe the experimentation you did:
\LeaveSpace[3in]{}
\item Describe the results:
\LeaveSpace[4in]{}

\item If you had more time, what next steps would you take?
\end{enumerate}

\end{document}